---
title: "Data 621 Homework 1"
author: "Critical Thinking Group 3"
date: "`r Sys.Date()`"
output:
  prettydoc::html_pretty:
    theme: cayman
    highlight: github
vignette: >
  %\VignetteIndexEntry{Vignette Title}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r include=FALSE}
library(tidyverse)

training_set<-read_csv("https://raw.githubusercontent.com/Vy4thewin/criticalthinking3/main/moneyball-training-data.csv")
test_set<-read_csv("https://raw.githubusercontent.com/Vy4thewin/criticalthinking3/main/moneyball-evaluation-data.csv")
````

### Exploring the MoneyBall game statistics from 1871 to 2006 [Made Up name for team- Gators? ]|DATA EXPLORATION 
### Preparing the game stats data set for regressions |DATA PREPARATION
### Selecting the predictor values for our model| BUILD MODELS 
### The final model's SELECT MODELs


[Internal notes:
Hi! Wanted to play around with the data set to get a feel for our next HW. If there are thoughts that may be included in the final version of the HW; It will not be surrounded by brackets [].-VH
]


### Appendix

```{r}
#Let's review our training set and if the training set needs updates before its process
summary(training_set)

#See the number of NAs per columns. Noticed Batters hit per pitch has the highest 
colSums(is.na(training_set))

#Visually checking for a positive linear trend with a singular predictor value
#It's positive but there's a large grouping around 500
g<-ggplot(training_set,aes(x=TEAM_BATTING_BB,y=TARGET_WINS))+geom_point()

#[It's not mentioned the way the data set was collected so we cannot fill NAs as zeros. We can remove all NAs which shrinks our data set to 191. Removing columns w/ Na may removed predictors with the best influence on the Wins of this team ]

#Noticed there are entries with NAs values that cannot be replaced with a formula. These entries will be removed from the training_set to prevent a inaccurate model
train.clean<-na.omit(training_set)


#checking the summary of our clean data set
summary(train.clean)

#[We can use the step() to find predictors instead as it's r version of the stepwise algorithm instead of the manually creating three models but we can also come up for the journey]


````



